An Adolescent Mini-Neptune in the Kepler Field

### 2021-07-02 paper/ms.tex
add Prot window slider and unc calibration

\begin{abstract}
  Kepler\,1627A is a G8V star previously found to host a
  $3.6\,R_\oplus$ mini-Neptune on a 7.2\,day orbit.  The star was
  selected for observation by the Kepler satellite because it is
  nearby ($d\approx 333\,{\rm pc}$) and resembles the Sun.  Combining
  data from Gaia with rotation periods from TESS, we show that
  Kepler\,1627 is a member of the $35\pm XX$\,Myr old $\delta$~Lyr
  cluster.  Our analysis advances Kepler\,1627Ab as the youngest
  planet with a precise age found by the main Kepler mission.
  Analyzing the photometry, we find evidence that starspot crossings
  {\it i)} distort the average transit shape, and {\it ii)} induce
  transit timing variations that suggest the planet is on a prograde
  orbit.  The possibility that star-planet interactions could induce
  flares in the system is also explored.  Future spectroscopic studies
  of \pn\ may help clarify the early evolution of the orbit and
  atmosphere for this mini-Neptune.  More generally, the merging of
  Gaia and TESS data offers an opportunity to significantly expand the
  census of age-dated planets; we provide a literature concatenation
  of young stars to support this effort.
\end{abstract}

@@

While thousands of exoplanets have been discovered orbiting nearby
stars, the vast majority of them are several billion years old.  This
results in difficulty testing theories for the origins of the
different families of planets, since many expected evolutionary
processes operate on timescales of less than 100 million years.

For instance, the ``mini-Neptune'' planets, thought to be made of
molten rocky cores and extended atmospheric envelopes of hydrogen and
helium, are expected to shrink in size by factors of several over
their first $10^8$ years \citep{owen_constraining_2020}.
Specifically, they start with sizes of 4--12\,$R_\oplus$ at the time
of disk dispersal ($\approx$$10^7$\,years), and shrink to sizes of
2--4\,$R_\oplus$ by 10$^8$ years.  This change is driven by a
combination of stellar irradiation and internal heat that powers an
outflow, which eventually depletes or entirely strips the envelope
from the rocky core \citep{Owen_Wu_2013,gupta_sculpting_2019}.
Discovering young planets and measuring their compositions and
atmospheric outflows are key steps toward testing this paradigm, which
is often invoked to explain the observed radius distribution of both
mature exoplanets \citep{Fulton_et_al_2017}, as well as younger
planets \citep{bouma_cluster_2020}.

@@

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.99\textwidth]{f2.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{
		{\bf Kepler and TESS views of the $\delta$\,Lyr cluster.} Colored
		points are kinematically selected members of the $\delta$\,Lyr
		cluster (black points in Figure~\ref{fig:XYZvtang}).  Both Kepler
		(gray squares) and TESS (colored points) observed portions of the
		cluster.  Naked-eye stars ($m_{\rm V}<6.5$) are shown in gray; four
		of them (orange points) have their names annotated.  Kepler\,1627 (green
		star) was observed during the entirety of the Kepler mission, and
		has been observed  by TESS for two lunar months to date.
		\label{fig:skychart}
	}
\end{figure*}

@@

		\subfloat{
			\includegraphics[width=0.87\textwidth]{f5.pdf}
		}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf The light curve of Kepler\,1627.}
    {\it Top}: 
		The full Kepler dataset spans 1{,}437 days (3.9 years), sampled at
		30 minute cadence.  A one hundred day segment is shown here.  The {\it top panel} shows the
		\texttt{PDCSAP} median-subtracted flux in units of
		parts-per-thousand ($\times 10^{-3}$).  The dominant signal
		is induced by starspots.  The model for the
    stellar variability (orange line) is subtracted below, revealing
    the transits of \pn, as well as other deviations from the stellar
    variability model.  The Figure Set available online shows the
    entire 3.9 years of available data.
    {\it Bottom}:
		Phase-folded transit of Kepler 1627b, with stellar
			variability removed.
    The 2-$\sigma$ model uncertainties and the maximum {\it a
    posteriori} model are shown as the faint purple band, and the dark
    purple line.  Gray points are individual \texttt{PDCSAP} flux
    measurements; black points are binned to 20 minutes, with a
    representative error bar shown.
    The in-transit residual is robust against detrending methods; we
    believe it is caused by starspot crossings (see text).
    \label{fig:lc}
  }
\end{figure*}

@@

Comparing our young star list against the Kepler field yielded two
discoveries.  The first, to be discussed in an upcoming analysis
by J. Curtis et al{.}, is that Kepler observed the $\approx$350\,Myr
open cluster Theia\,520 (UBC\,1).  At least five Kepler planets are
known in the cluster, including the Kepler-52 and Kepler-968 systems.
The second discovery, and the focus of this work, is that Kepler
observed a portion of the $\delta$\,Lyr cluster, shown in
Figure~\ref{fig:skychart}.

While we have yet to perform an in-depth planet search of the {\bf
XXX} light curves Kepler acquired for the \cn, there is already a
known planet in the cluster: Kepler\,1627b.  Given the
planet's previous statistical validation by \citet{morton_false_2016},
we decided to examine the age of the star and its cluster more closely.
We collected the previously measured {\bf XXX} Kepler
rotation periods (CITEP MCQUILLAN2014), and measured {\bf YYY} new stellar
rotation periods for cluster members observed by TESS.
We also performed an isochronal analysis.
The results are shown in Figure~\ref{fig:age}.
From the ensemble perspective, which provides the strongest
constraints (SEE CITET SODERBLOM 2014),
the age of the star is $XX \pm YY$\,Myr.
Individual measurements of the star's rotation period and lithium
abundance corroborate this conclusion.

If the Kepler\,1627b transit signal is created by a bonafide planet,
this age would make it the youngest planet yet found by the main
Kepler mission.  A portion of the light curve is shown in
Figure~\ref{fig:lightcurve}, and the phase-folded transit is shown in
Figure~\ref{fig:phasefold}.  The dominant signal in the
\texttt{PDCSAP} photometry is a quasi-period starspot signal with a
peak-to-peak amplitude that varies between roughly 2\% and 8\%,
depending on the degree of asymmetry in spot coverage between the
stellar hemispheres.  The starspot signal repeats roughly every
2.606\,days, though this value varies by a few percent depending on
which portion of the light curve is analyzed {\bf (FIXME: do!)}.  The
details of how we model this signal to derive the parameters of the
planet are discussed in Section~\ref{app:gptransit}.  The results,
given in Table~\ref{tab:posterior}, are consistent with a mini-Neptune
sized planet ($3.56\pm 0.04\,R_\oplus$) on a circular 7.2\,day orbit,
around a G8V star ($0.91 \pm 0.05 R_\odot$) {\bf fixme: age}.

>>>>

\begin{abstract}
  Kepler\,1627A is a G8V star previously found to host a
  $3.6\,R_\oplus$ mini-Neptune on a 7.2\,day orbit.  The star was
  selected for observation by the Kepler satellite because it is
  nearby ($d\approx 333\,{\rm pc}$) and resembles the Sun.  Combining
  data from Gaia with rotation periods from TESS, we show that
  Kepler\,1627 is a member of the $35\pm XX$\,Myr old $\delta$~Lyr
  cluster.  Kepler\,1627Ab is therefore the youngest planet with a
  precise age found so far by the main Kepler mission.
  Analyzing the photometry, we find evidence that starspot crossings
  distort the average transit shape, and induce transit timing
  variations that suggest the planet is on a prograde orbit.  The
  possibility that star-planet interactions could induce
  flares in the system is also explored.  Future spectroscopic studies
  of \pn\ may help clarify how the orbit and atmosphere of this
  mini-Neptune evolve.  More generally, the combination of Gaia and TESS
  offers an opportunity to significantly expand the census of
  age-dated planets; we provide a literature concatenation of young
  stars to support this effort.
\end{abstract}

@@

While thousands of exoplanets have been discovered orbiting nearby
stars, the vast majority of them are several billion years old.  This
makes it difficult to test theories for the origins of the
different families of planets, since many expected evolutionary
processes operate on timescales of less than 100 million years.

For instance, the ``mini-Neptune'' planets, thought to be made of
molten rocky cores and extended atmospheric envelopes of hydrogen and
helium, are expected to shrink in size by factors of several over
their first $10^8$ years \citep{owen_constraining_2020}.
Specifically, they start with sizes of 4--12\,$R_\oplus$ at the time
of disk dispersal ($\lesssim$$10^7$\,years), and shrink to sizes of
2--4\,$R_\oplus$ by 10$^8$ years.  This change is driven by a
combination of stellar irradiation and internal heat that powers an
outflow, which eventually depletes or entirely strips the envelope
from the rocky core \citep{Owen_Wu_2013,gupta_sculpting_2019}.
Discovering young planets and measuring their compositions and
atmospheric outflows are key steps toward testing this paradigm, which
is often invoked to explain the observed radius distribution of both
mature exoplanets \citep{Fulton_et_al_2017}, as well as younger
planets \citep{bouma_cluster_2020}.

@@

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.99\textwidth]{f2.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{
		{\bf Kepler and TESS views of the $\delta$\,Lyr cluster.} Colored
		points are kinematically selected members of the $\delta$\,Lyr
		cluster.  Both Kepler
		(gray squares) and TESS (colored points) observed portions of the
		cluster.  Naked-eye stars ($m_{\rm V}<6.5$) are shown in gray; four
		of them (orange points) have their names annotated.  Kepler\,1627 (green
		star) was observed during the entirety of the Kepler mission, and
		has been observed  by TESS for two lunar months to date.
		\label{fig:skychart}
	}
\end{figure*}

@@

		\subfloat{
			\includegraphics[width=0.87\textwidth]{f5.pdf}
		}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf The light curve of Kepler\,1627.}
    {\it Top}: 
		The full Kepler dataset spans 1{,}437 days (3.9 years), sampled at
		30 minute cadence.  A one hundred day segment is shown here.  The {\it top panel} shows the
		\texttt{PDCSAP} median-subtracted flux in units of
		parts-per-thousand ($\times 10^{-3}$).  The dominant signal
		is induced by starspots.  The model for the
    stellar variability (orange line) is subtracted below, revealing
    the transits of \pn, as well as other deviations from the stellar
    variability model.  The Figure Set available online shows the
    entire 3.9 years of available data.
    {\it Bottom}:
		Phase-folded transit of Kepler 1627b, with stellar
			variability removed.
    The 2-$\sigma$ model uncertainties and the maximum {\it a
    posteriori} model are shown as the faint purple band, and the dark
    purple line.  Gray points are individual \texttt{PDCSAP} flux
    measurements; black points are binned to 20 minutes, with a
    representative error bar shown.
    The asymmetric residual visible during the transit is robust
    against detrending methods; we believe it is caused by starspot
    crossings.
    \label{fig:lc}
  }
\end{figure*}

@@

Comparing our young star list against the Kepler field yielded two
discoveries.  The first, to be discussed in an upcoming analysis
by J. Curtis et al{.}, is that Kepler observed the $\approx$350\,Myr
open cluster Theia\,520 (UBC\,1).  At least six Kepler planets are
known in the cluster, including the Kepler-52 and Kepler-968 systems.
The second discovery, and the focus of this work, is that Kepler
observed a portion of the $\delta$\,Lyr cluster, shown in
Figure~\ref{fig:skychart}.

Kepler\,1627 (KIC 6184894; KOI 5245; TIC 120105470) was previosuly
reported to host a close-in mini-Neptune planet candidate,
Kepler\,1627Ab
\citep{2012ApJS..199...24T,thompson_planetary_2018}
Given the
planet's previous statistical validation by \citet{morton_false_2016},
we decided to examine the age of the star and its cluster more closely.
We collected the previously measured {\bf XXX} Kepler
rotation periods (CITEP MCQUILLAN2014), and measured {\bf YYY} new stellar
rotation periods for cluster members observed by TESS.
We also performed an isochronal analysis.
The results are shown in Figure~\ref{fig:age}.
From the ensemble perspective, which provides the strongest
constraints (SEE CITET SODERBLOM 2014),
the age of the star is $XX \pm YY$\,Myr.
Individual measurements of the star's rotation period and lithium
abundance corroborate this conclusion.

% rotation period from plot_rotation_period_windowslider.py

If the \pn\ transit signal is created by a bonafide planet, this age
would make it the youngest planet yet found by the main Kepler
mission\footnote{The re-purposed K2 mission found two slightly younger
systems: K2-33b \citep{David_et_al_2017,Mann_K2_33b_2016} and V1298
Tau \citep{david_four_2019}}.  A portion of the light curve is shown
in Figure~\ref{fig:lightcurve}, and the phase-folded transit is shown
in Figure~\ref{fig:phasefold}.  The dominant signal in the
\texttt{PDCSAP} photometry is a quasi-period starspot signal with a
peak-to-peak amplitude that varies between roughly 2\% and 8\%,
depending on the degree of asymmetry in spot coverage between the
stellar hemispheres.  The starspot signal repeats every
$2.642\pm0.042$\,days, where we have adopted an uncertainty based on
the scatter observed between different windows (quarters) of the
time-series data.  The details of how we model this starspot signal to
derive the parameters of the planet are discussed in
Section~\ref{app:gptransit}.  The results, given in
Table~\ref{tab:posterior}, are consistent with a mini-Neptune sized
planet ($3.56\pm 0.04\,R_\oplus$) on a circular 7.2\,day orbit
($e<0.XX$), around
a G8V star ($0.91 \pm 0.05 R_\odot$). 
{\bf TODO quote the eccentricity limit}
